Motional Quantum Ground-State Cooling Outside the Lamb-Dicke Regime -- Supplemental material

### 2017-07-20 paper.tex
Implement all obvious suggestions from Lee and Jessie

@@

Our experiment begins by loading a single sodium atom into an optical tweezer from a MOT
for $250$ ms~\cite{Hutzler2017-LightShifts}, and has an overall repetition rate of $2.5$ Hz.
The tweezer is created by focusing a $700$ nm laser beam through an NA=$0.55$ objective to
create a focused beam with waist $0.7 \mu$m.
With $45$ mW of tweezer power, the trapping frequencies are
$\{\omega_1,\omega_2,\omega_3\}/2\pi = \{69(1), 430(4), 590(5)\}\ \text{\text{kHz}}$.
After attempting to load an atom from the MOT, an image is taken with $1.5$ ms exposure time to determine if it was a success.
The same imaging light is used for polarization gradient cooling (PGC), which
reduces the temperature of the single atom to $70$ $\mu$K,
corresponding to an initial mean motional state along the three different axes of
$\{\bar n_1, \bar n_2, \bar n_3\}=\{21(2),\, 3.4(2),\, 2.5(2)\}$, and therefore initial 3D ground state probability of $P_0=0.3$\%.
A beam with $\sigma^-$-polarization is used to initialize the atom into
the $|F=2, m_F=-2\rangle$ stretched state via optical pumping.
The hyperfine state detection is performed by pushing out atoms in the stretched state with a strong
beam on resonance with the $|F=2, m_F=-2\rangle$ to $|F'=3, m_{F'}=-3\rangle$ transition before
taking the second image to determine if the atom was pushed out.
The state preparation and detection have a combined fidelity of larger than $99$\%.

@@

It is important that the re-pumping step scatter as few photons as possible to minimize heating.  Therefore, we use a $\sigma^-$-polarized laser resonant with the D1 line since the $|F=2, m_F=-2\rangle$ is a dark state in this case \fxnote{cite ion and K39}. Compared to using a D2 repump beam, from which the $|F=2, m_F=-2\rangle$ state could always scatter a photon via the excited $|F=3, m_F=-3\rangle$ state, we find a reduction in the scattering rate of a factor of $130(20)$.

@@

Guided by the simulation results,
we construct our axial cooling sequence by starting at the two highest
observed cooling sidebands (8th- and 7th-orders)
and decreasing the orders after $6$ to $15$ cycles.
This process is repeated until we reach the first order cooling sideband, and most of the
population is in the first few excited states.  We then switch to cooling only on the
first order sideband with two different pulse lengths in order to efficiently cool atoms in the
few remaining motional states.
The radial cooling is performed similarly to the axial cooling with the initial cooling orders being
the 2nd and 1st order and switching to first order only after $20$ to $30$ pulses.
This sequence gives good initial cooling performance, which can then be used to measure experimental
parameters, including the Rabi rates of the Raman beams and optical pumping rates,
in order to optimize the sequence further.

@@

Our final cooling results are shown in Fig.~\ref{f-radial} and \ref{f-axial}.
In total, $1000$ cooling pulses (with a total duration of $100$ ms) are applied
along three axes with cooling beginning on the radial second order.
To characterize the single atom thermal occupation before and after cooling,
we perform Raman sideband thermometry.
For the more tightly confined (non-degenerate) radial directions,
we observe clear first-order heating, first-order cooling,
and second-order cooling sidebands before RSC as is shown in red in Figure \ref{f-radial}A.
After cooling, the first and second order cooling sidebands on both radial axes are suppressed.
It is important to note that the simple sideband thermometry formula,
which predicts the ratio of the cooling and heating sideband heights
to be $\bar n / (\bar n + 1)$~\cite{Monroe1995},
assumes that the coupling strength for different motional states to be proportional to $\sqrt{n}$.
However, outside of the LD regime,
the coupling strength rapidly deviates from this simple scaling rule,
as shown in figure \ref{f-ld}A.  Since the Raman spectroscopy is performed
with relatively large Raman LD parameters $\eta^R$, we cannot use this formula to
measure $\bar n$.
Instead, given the absence of the second order cooling sideband,
we assume \fxnote{[What is the cutoff for the validity of this assumption, and are we below that cutoff?]} that the atoms are cold enough to have stronger coupling to the first order cooling
sideband than the second order, in which case the lower bound of the ground state population
can be computed from the heights of the sidebands by assuming all of the excitations are in the
$n=1$ state. In doing this, we found a ground state population of $90(2)$\%
and $94(3)$\% along the two radial directions.
We also use an independent measurement of Rabi flopping on the carrier (Fig. \ref{f-radial}B)
and first-order heating(Fig. \ref{f-radial}C) sidebands
to verify the ground state population~\cite{Meekhof1996}.
The data is fitted using the same Monte-Carlo simulation we used to simulate the cooling
which yields $89(2)$\% and $92(2)$\% and show good agreement with
the population extracted from the sideband thermometry.
We also extracted the initial temperature before cooling
from the motional-state decohered Rabi flopping to obtain the initial temperature of $70 \mu$K.

@@

Combining the axial and radial cooling results,
we have prepared a single Na atom with 3D ground state probability of $77(4)$\%.
We measuring a heating rate of the single atoms in the tweezer to be $0.9$\%/s
due to off-resonant scattering of the trapping light\cite{Grimm2000}.
The ground state preparation fidelity is currently limited by off-resonant scattering
from the Raman beams,
which are measured to be between $3$ to $15$ kHz for the detuning and power we use,
and the resonance shift caused by magnetic field fluctuations.

@@

% Lamb-Dicke parameter square explain
% why not just molecules instead of polar molecules
% entropy
% use different line shape for rabi flopping
% mention axis definitions
% heating from trap -> decrease of ground state procetage (maybe add temprature too)
% at the same time, the atom is cooled via PGC instead of "the same imaging light is used to"
% 1 sentence to explain why anisotropy?
% Make sure the caption for the before cooling axial rabi flopping mentions decoherence
% Mention after cooling is non-thermal?
% Make it clear that we cool for 6-15 cycles each time before going to the next set of orders in axial

% sm:
% mention the simulation results are 3d
% effect of changing groups.
% mention effects that are included first.

>>>>

    The asymptotic values for the red lines are low also because of resonance fluctuation.

@@

Our experiment begins by loading a single sodium atom into an optical tweezer from a MOT
for $250$ ms~\cite{Hutzler2017-LightShifts}, and has an overall repetition rate of $2.5$ Hz.
The tweezer is created by focusing a $700$ nm laser beam through an NA=$0.55$ objective to
create a focused beam with waist $0.7 \mu$m.
With $45$ mW of tweezer power, the trapping frequencies along the three axes are
$\{\omega_1,\omega_2,\omega_3\}/2\pi = \{69(1), 430(4), 590(5)\}\ \text{\text{kHz}}$
where axis 1 is along the weakly confined axial direction and axis 2 and 3 are along the
tightly confined radial directions.
After attempting to load an atom from the MOT, an image is taken with $1.5$ ms exposure time to determine if it was a success.
At the same time, the atom is cooled via polarization gradient cooling (PGC),
which reduces the temperature of the single atom to $70$ $\mu$K,
corresponding to an initial mean motional state along the three different axes of
$\{\bar n_1, \bar n_2, \bar n_3\}=\{21(2),\, 3.4(2),\, 2.5(2)\}$, and therefore initial 3D ground state probability of $P_0=0.3$\%.
A beam with $\sigma^-$-polarization is used to initialize the atom into
the $|F=2, m_F=-2\rangle$ stretched state via optical pumping.
The hyperfine state detection is performed by pushing out atoms in the stretched state with a strong
beam on resonance with the $|F=2, m_F=-2\rangle$ to $|F'=3, m_{F'}=-3\rangle$ transition before
taking the second image to determine if the atom was pushed out.
The state preparation and detection have a combined fidelity of larger than $99$\%.

@@

It is important that the re-pumping step scatter as few photons as possible to minimize heating.
Therefore, we use a $\sigma^-$-polarized laser resonant with the D1 line
since $|F=2, m_F=-2\rangle$ is a dark state in this case \fxnote{cite ion and K39}.
Compared to using a D2 repump beam, from which the $|F=2, m_F=-2\rangle$ state
could always scatter a photon via the excited $|F=3, m_F=-3\rangle$ state,
we find a reduction in the scattering rate of a factor of $130(20)$.

@@

Guided by the simulation results,
we construct our axial cooling sequence by starting at the two highest
observed cooling sidebands (8th- and 7th-orders)
and decreasing the orders to the next pair after every $6$ to $15$ cycles.
This process is repeated until we reach the first order cooling sideband, and most of the
population is in the first few excited states. We then switch to cooling only on the
first order sideband with two different pulse lengths in order to efficiently cool atoms in the
few remaining motional states.
The radial cooling is performed similarly to the axial cooling with the initial cooling orders being
the 2nd and 1st order and switching to first order only after $20$ to $30$ pulses.
This sequence gives good initial cooling performance, which can then be used to measure experimental
parameters, including the Rabi rates of the Raman beams and optical pumping rates,
in order to optimize the sequence further.

@@

Our final cooling results are shown in Fig.~\ref{f-radial} and \ref{f-axial}.
In total, $1000$ cooling pulses (with a total duration of $100$ ms) are applied
along three axes with cooling beginning on the radial second order.
To characterize the single atom thermal occupation before and after cooling,
we perform Raman sideband thermometry.
For the more tightly confined (non-degenerate) radial directions,
we observe clear first-order heating, first-order cooling,
and second-order cooling sidebands before RSC as is shown in red in Figure \ref{f-radial}A.
After cooling, the first and second order cooling sidebands on both radial axes are suppressed.
It is important to note that the simple sideband thermometry formula,
which predicts the ratio of the cooling and heating sideband heights
to be $\bar n / (\bar n + 1)$~\cite{Monroe1995},
assumes that the coupling strength for different motional states to be proportional to $\sqrt{n}$.
However, outside of the LD regime,
the coupling strength rapidly deviates from this simple scaling rule,
as shown in figure \ref{f-ld}A. Since the Raman spectroscopy is performed
with relatively large Raman LD parameters $\eta^R$, we cannot use this formula to
measure $\bar n$.
Instead, given the absence of the second order cooling sideband,
we assume \fxnote{[What is the cutoff for the validity of this assumption, and are we below that cutoff?]} that the atoms are cold enough to have stronger coupling to the first order cooling
sideband than the second order, in which case the lower bound of the ground state population
can be computed from the heights of the sidebands by assuming all of the excitations are in the
$n=1$ state. In doing this, we found a ground state population of $90(2)$\%
and $94(3)$\% along the two radial directions.
We also use an independent measurement of Rabi flopping on the carrier (Fig. \ref{f-radial}B)
and first-order heating(Fig. \ref{f-radial}C) sidebands
to verify our assumption about the distribution and the ground state population~\cite{Meekhof1996}.
The data is fitted using the same Monte-Carlo simulation we used to simulate the cooling
which yields $89(2)$\% and $92(2)$\% and show good agreement with
the population extracted from the sideband thermometry.
We also extracted the initial temperature before cooling
from the motional-state decohered Rabi flopping to obtain the initial temperature of $70 \mu$K.

@@

Combining the axial and radial cooling results,
we have prepared a single Na atom with 3D ground state probability of $77(4)$\%.
We measured a heating rate of the single atoms in the tweezer to be $67$nK/ms
due to off-resonant scattering of the trapping light, which
corresponds to $0.9$\%/ms decrease in ground state population\cite{Grimm2000}.
The ground state preparation fidelity is currently limited by off-resonant scattering
from the Raman beams,
which are measured to be between $3$ to $15$ kHz for the detuning and power we use,
and the resonance shift caused by magnetic field fluctuations.

@@

% Lamb-Dicke parameter square explain
% why not just molecules instead of polar molecules
% use entropy to describe the effect of large \eta_{eff}
% 1 sentence to explain why anisotropy?
% Mention after cooling is non-thermal?

### 2017-07-20 sm.tex
Implement all obvious suggestions from Lee and Jessie

\begin{enumerate}
\item \emph{All Raman pulses resonantly drive only one sideband/carrier order.}  We tweak the pulse shape, power and length as discussed in the text, for example by using Blackman pulses, to reduce coupling to other orders. We do see deviation from this approximation in the experiment, though mostly in the long detection pulses, and it is less of a concern for the short cooling pulses.
\item \emph{No heating from other sources.}\\
  We measured the heating rate and found that it results mainly from off-resonant scattering
  of photons, which is included in the simulation.  In particular, we can ignore heating from the switching trap.
\item \emph{Same trapping frequency for all states.}\\
  This effect is negligible for Na in the ground state due to the small fine structure splitting in the excited state.
  It is easy to include this effect if the calculation needs to be repeated on a system where
  this is not the case (e.g. directly laser cooled molecules)
\end{enumerate}

@@

\begin{figure}[t]
  \includegraphics[width=8.5cm]{imgs/simcool_no_scatter.pdf}
  \caption{Simulated ground state population as a function of cooling cycles without
    technical imperfections. The ground state probability reaches $>99\%$. \label{f-no-scatter}}
\end{figure}
\begin{figure}[t]
  \includegraphics[width=8.5cm]{imgs/simcool_real.pdf}
  \caption{Simulated ground state population as a function of cooling cycles including effect of
    off-resonant scattering from all beams and OP misalignment.
    For this particular sequence, the ground state probability saturates to $\approx88\%$.
    \label{f-real}}
\end{figure}

If off-resonant scattering and optical pumping polarization misalignment are ignored,
the simulation shows that there isn't a cooling limit
despite the large Lamb-Dicke parameter and high initial temperature (Fig \ref{f-no-scatter}).
However, when these effects are taken into account, the ground state probability can saturate at
a lower value (Fig \ref{f-real}) and require much more careful tweaking.

@@

>>>>

\begin{enumerate}
\item \emph{All Raman pulses resonantly drive only one sideband/carrier order.}  We tweak the pulse shape, power and length as discussed in the text, for example by using Blackman pulses, to reduce coupling to other orders. We do see deviation from this approximation in the experiment, though mostly in the long detection pulses, and it is less of a concern for the short cooling pulses.
\item \emph{No heating from other sources.}\\
  We measured the heating rate and found that it results mainly from off-resonant scattering
  of photons, which is included in the simulation.  In particular, we can ignore heating from the switching trap.
\item \emph{Same trapping frequency for all states.}\\
  This effect is negligible for Na in the ground state due to the small fine structure splitting in the excited state.
  It is straightforward to include this effect if the calculation needs to be repeated
  on a system where this is not the case (e.g. directly laser cooled molecules).
\end{enumerate}

@@

\begin{figure}[t]
  \includegraphics[width=8.5cm]{imgs/simcool_no_scatter.pdf}
  \caption{Simulated 3D ground state population as a function of cooling cycles without
    technical imperfections. The ground state probability reaches $>99\%$. \label{f-no-scatter}}
\end{figure}
\begin{figure}[t]
  \includegraphics[width=8.5cm]{imgs/simcool_real.pdf}
  \caption{Simulated 3D ground state population as a function of cooling cycles including effect of
    off-resonant scattering from all beams and OP misalignment.
    For this particular sequence, the ground state probability saturates to $\approx88\%$.
    \label{f-real}}
\end{figure}

If off-resonant scattering and optical pumping polarization misalignment are ignored,
the simulation shows that there is not a cooling limit
despite the large Lamb-Dicke parameter and high initial temperature (Fig \ref{f-no-scatter}).
However, when these effects are taken into account, the ground state probability can saturate at
a lower value (Fig \ref{f-real}) and require much more careful tweaking.
The change of slop in these plots corresponds to changing the sideband orders used for cooling.

@@

% mention effects that are included first?
